---
title: "White Wine Analysis"
author: "Matthew Bonilla"
date: "1/17/2018"
output: html_document
---


White Wine Analysis by Matthew Bonilla
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(GGally)
library(scales)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wines <- read.csv("WineQualityWhites.csv")
```

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

This report will analyze the quality of white wines and how chemical factors, 
such as acidity, sugar, pH levels, and alcohol content affect it. There are 11 
input variables taken from physiochemical tests that make up one output data 
based on sensory data with a dataset of about 4,900 wines. 

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
dim(wines)

summary(wines)
```

The dataset contains 4,898 observations with 12 variables. (The first, X, is 
simply the count of the observation).

```{r echo = FALSE, Univaraite_Plots}
ggplot(wines, aes(quality)) + geom_bar()

```

Taking a look at our output variable, quality, we can see that we do not have 
as much points less than 5 and more than 7. 

```{r}
ggplot(wines, aes(alcohol)) + geom_histogram(binwidth = .3)
```

From the input variables present, alcohol content, as a percentage, is one that 
most are familiar with. We can see that most of the alcohol percentage of the 
wines are 10 and below. 

```{r}
ggplot(wines, aes(residual.sugar)) + geom_histogram()

ggplot(wines, aes(residual.sugar)) + geom_histogram(binwidth = .5)
```

It looks as if most wines have residual sugar levels below 20 gram/liter. We do
have a few above 20 and one above 60 gram/liter. 

```{r}
ggplot(wines, aes(fixed.acidity)) + geom_histogram(binwidth = .2)

ggplot(wines, aes(volatile.acidity)) + geom_histogram(binwidth = .02)

ggplot(wines, aes(citric.acid)) + geom_histogram(binwidth = .03)
```

Fixed acidity, volatile acidity, and citric acid all have a bell curve like 
structure with all three having what look to be outleirs on the higher end. 
There seems to be a spike in the amount of wines with a citric acid of 0.5. 

```{r}
summary(wines$citric.acid)

table(wines$citric.acid)
```

This would definitely be something to look at in the future, why is there such 
a large amount of wines with citric acid of 0.49 yet levels of 0.48 and 0.5 are 
not that high. 

```{r}
ggplot(wines, aes(chlorides)) + geom_histogram(binwidth = 0.003)

ggplot(wines, aes(chlorides)) + 
  geom_freqpoly(binwidth = 0.003) + 
  scale_y_continuous(trans = log10_trans())
```

We see that a large concentration of chloride content lay between 0.025 and 
0.075. When we take the log of the counts, we see a good amount of wines with chlorides above 0.1. 

```{r}
ggplot(wines, aes(free.sulfur.dioxide)) + geom_histogram(binwidth = 5) + 
  geom_freqpoly(binwidth = 5) + 
  coord_cartesian(xlim = c(0, 150))

ggplot(wines, aes(total.sulfur.dioxide)) + geom_histogram(binwidth = 5) + 
  geom_freqpoly(binwidth = 5)

p1 <- ggplot(wines, aes(free.sulfur.dioxide)) + 
  geom_histogram(binwidth = 5, alpha = .5) +
  coord_cartesian(xlim = c(0, 450))
p2 <- ggplot(wines, aes(total.sulfur.dioxide)) + 
  geom_histogram(binwidth = 5, alpha = .5) + 
  coord_cartesian(xlim = c(0, 450))
grid.arrange(p1, p2)

fsd <- data.frame(xx = wines$free.sulfur.dioxide, y = 'free.sulfur.dioxide')
tsd <- data.frame(xx = wines$total.sulfur.dioxide, y = 'free.sulfur.dioxide')
suldio <- rbind(fsd, tsd)

ggplot(suldio,aes(x = xx)) + 
  geom_histogram(data =fsd,fill = 'blue',
                 alpha =0.5, binwidth = 5) +
  geom_histogram(data = tsd, fill = 'red',
                 alpha = 0.5, binwidth = 5) + 
  coord_cartesian(xlim = c(0, 450))

summary(wines$free.sulfur.dioxide)

## https://stackoverflow.com/questions/6957549/overlaying-histograms-with-ggplot2-in-r
```

## FIND OUT HHOW TO COMBINE GRAPHS INTO ONE AND OVERLAY ON TOP OF EACH OTHER. 
##THIS WILL MAKE THE GRAPH LOOK COOL AS HECK. 

```{r}
ggplot(wines, aes(x = sulphates)) + 
  geom_histogram(binwidth = 0.02)

summary(wines$sulphates)
```

### WRITE SOMETHING WITTY HERE PLEASE
### FOR THE LOVE OF FOOD PLEASE

```{r}
ggplot(wines, aes(x = pH)) + geom_histogram(binwidth = 0.05)
```

Pretty normal distribution around 3.15

```{r}
ggplot(wines, aes(x = density)) + geom_histogram(binwidth = 0.001)
```

Seem to be distributed around .993


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
dir


### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!